Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-04-22 ms.tex
Adding time dependent derivation up to time-integrated spectra vs. steady state

There are certain values of $\omega$ for which a resonant response is produced. We will call these eigenfrequencies $\omega_{nm}$. Near these frequencies, the solution will have the form

\be
J(n,\sigma,s) & \simeq \frac{ J_{nm}(\sigma) }{s - s_{nm}} + {\rm terms\ which\ vary\ slowly\ in\ s}.
\ee

@@

\be
J(r,\sigma,t) & = &  j_0(\kappa_n r)  \int \frac{d\omega}{2\pi} e^{-i\omega t}J(n,\sigma,\omega)
\nonumber \\ & \simeq & j_0(\kappa_n r)  \int \frac{d\omega}{2\pi} e^{-i\omega t} \frac{ J_{nm}(\sigma) }{\omega - \omega_{nm}} 
\nonumber \\ & \simeq & j_0(\kappa_n r)  J_{nm}(\sigma) e^{-i\omega_{nm}t}.
\ee

To find this result, we have closed the integral along the $i\omega$ axis by a semi-circle at infinity, and we then used the residue theorem to evaluate the answer. The contribution from the semi-circle at infinity is zero. Summing over all spatial n's and over all eigenmodes $m$ for a given $n$, we get the final form of the result
\be
J(r,\sigma,t) & = & \sum_{n=1}^\infty j_0(\kappa_n r)  \sum_m J_{nm}(\sigma) e^{-i\omega_{nm}t}.
\ee
Using the derivative
\be
\frac{dj_0(\kappa_n r)}{dr} \rfloor_R & =& \frac{d}{dr}  \rfloor_R \left( \frac{\sin(\kappa_n r)}{\kappa_n r} \right)
=  \left( \frac{\cos(\kappa_n R)}{R} - \frac{\sin(\kappa_n R)}{\kappa_n R^2} \right)\rfloor_R
\nonumber \\ & = & \frac{(-1)^n}{R},
\ee
the flux at the surface is
\be
F(R,\sigma,t) & =& - \frac{4\pi}{3k\phi} \frac{dJ(R,\sigma,t)}{dr} 
= - \frac{4\pi}{3k\phi R}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-i\omega_{nm}t}.
\ee
Multiplying by $4\pi R^2$ gives the energy per time per frequency
\be
\frac{dE}{dtd\nu} & = & - \frac{16\pi^2 R}{3k\phi}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-i\omega_{nm}t}.
\label{eq:dEdtdnu}
\ee
If we integrate this expression over time, giving a factor $1/s_{nm}$, and over $d\nu = \sqrt{3/2} \Delta^2 \phi d\sigma$, a ``sum rule" results
\be
1 & = &  \sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3kE} \sum_{nm} (-1)^n (-i\omega_{nm})^{-1} \int d\sigma J_{nm}(\sigma).
\ee
This non-trivial expression provides a check on the values of $s_{nm}$ and the functions $J_{nm}(\sigma)$. This expression can also be written as
\be
1 & =& \sum_{nm} P_{nm},
\label{eq:sumrule}
\ee
where the contribution of each eigenfunction is
\be
P_{nm} & \equiv & -\sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3kE}  (-1)^n (i\omega_{nm})^{-1} \int d\sigma J_{nm}(\sigma).
\ee
These coefficients can have either sign depending on the value of $n$.

Integrating $dE/dt d\nu$ over all times should give a distribution for the emitted frequencies, and this should agree with the steady state solution. Dividing by the energy $E$, we find the probability distribution
\be \label{eq:spectrum}
P(\nu) & = &  \frac{16\pi^2 R}{3k\phi E}  \sum_{nm} (-1)^n s_{nm}^{-1} J_{nm}(\sigma).
\ee
Integrating over $\nu$ gives unity if the sum rule in Equation \ref{eq:sumrule} is true.

It is expected that the time-integrated spectrum of the response to an impulse will agree well with the solution for the steady-state spectrum (Eq. \ref{eq:spectrum}). We first evaluate the time-dependent response using our eigenfrequency sweeping method. Once the probabilities $P_{nm}(\sigma)$ are calculated as above, the escape time distributions for all photon frequencies are calculated by summing over the frequency eigenfunctions $P_{nm} \exp(s_{nm}t)$. Integrating $P_{nm}$ with respect to $t$, we obtain

>>>>

Let us define the damping time to be $\gamma \equiv i\omega$, which is real-valued and positive for damped solutions. There are certain values of $\gamma$ for which a resonant response is produced which we will call $\gamma_{nm}$. Here, $m$ is an index that represents individual temporal eigenmodes, as opposed to $n$ which represents the spatial mode number. Near these frequencies, the solution will have the form

\be
J(n,\sigma,\gamma) & \simeq \frac{ J_{nm}(\sigma) }{\gamma - \gamma_{nm}} + {\rm terms\ which\ vary\ slowly\ in\ \gamma}.
\ee

@@

\be
J(r,\sigma,t) & = &  j_0(\kappa_n r)  \int \frac{d\omega}{2\pi} e^{-\gamma t}J(n,\sigma,\gamma)
\nonumber \\ & \simeq & j_0(\kappa_n r)  \int \frac{d\gamma}{2\pi} e^{-\gamma t} \frac{ J_{nm}(\sigma) }{\gamma - \gamma_{nm}} 
\nonumber \\ & \simeq & j_0(\kappa_n r)  J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee

(TODO: Rewrite) Here we have closed the integral along the $\gamma$ axis by a semi-circle at infinity, using the residue theorem to obtain the answer. The contribution from the semi-circle at infinity is zero. Summing over all spatial eigenmodes $n$ and over all temporal eigenmodes $m$ for a given $n$, we get the final form of the result
\be
J(r,\sigma,t) & = & \sum_{n=1}^\infty j_0(\kappa_n r)  \sum_m J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
Using the derivative
\be
\frac{dj_0(\kappa_n r)}{dr} \rfloor_R & =& \frac{d}{dr}  \rfloor_R \left( \frac{\sin(\kappa_n r)}{\kappa_n r} \right)
=  \left( \frac{\cos(\kappa_n R)}{R} - \frac{\sin(\kappa_n R)}{\kappa_n R^2} \right)\rfloor_R
\nonumber \\ & = & \frac{(-1)^n}{R},
\ee
the flux at the surface is
\be
F(R,\sigma,t) & =& - \frac{4\pi}{3k\phi} \frac{dJ(R,\sigma,t)}{dr} 
= - \frac{4\pi}{3k\phi R}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
Multiplying by $4\pi R^2$ gives the energy per time per frequency
\be
\frac{dE}{dtd\nu} & = & - \frac{16\pi^2 R}{3k\phi}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\label{eq:dEdtdnu}
\ee
Integrating this over time yields a factor $-1/\gamma_{nm}$, and over $d\nu = \sqrt{3/2} \Delta^2 \phi d\sigma$, a ``sum rule" results
\be
1 & = &  \sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3kE} \sum_{nm} (-1)^n \gamma_{nm}^{-1} \int d\sigma J_{nm}(\sigma).
\ee
This non-trivial expression provides a check on the values of $\gamma_{nm}$ and the functions $J_{nm}(\sigma)$. This expression can also be written as
\be
1 & =& \sum_{nm} P_{nm},
\label{eq:sumrule}
\ee
where the contribution of each eigenfunction is
\be
P_{nm} & \equiv & \sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3kE}  (-1)^n \gamma_{nm}^{-1} \int d\sigma J_{nm}(\sigma).
\ee
These coefficients can have either sign depending on the value of $n$.

Integrating $dE/dt d\nu$ over all times should give a distribution for the emitted frequencies. It is expected that this time-integrated spectrum of the response to an impulse will agree well with the solution for the steady-state spectrum (Eq. \ref{eq:spectrum}). Dividing by the energy $E$, we find the probability distribution
\be \label{eq:spectrum}
P(\nu) & = &  -\frac{16\pi^2 R}{3k\phi E}  \sum_{nm} (-1)^n \gamma_{nm}^{-1} J_{nm}(\sigma).
\ee
Integrating over $\nu$ gives unity if the sum rule in Equation \ref{eq:sumrule} is true.
